A bird’s-eye view on the habitability of exoplanets via statistical learning techniques

### 2020-10-28 Discussion/Presentation.tex
PUB

\documentclass[compress]{beamer}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\author{Marzio De Corato}
\usepackage{varwidth}
\usepackage{beamerthemesplit}

@@

\title[Exoplanets survey via SL techniques]{A bird’s-eye view on the habitability of exoplanets via statistical learning techniques}

\author{Marzio De Corato}

@@

@@

\frame{\titlepage}

@@

\begin{frame}
\frametitle{Theoretical background - Planet features }
\begin{itemize}
\item\textbf{Distance}: in this work the mean planet distance from the host star (P\_D), the periastron (P\_PN) and the apastron (P\_A) as well the termal effective distance (P\_D\_E) from the host star  were considered. These quantities constrain the planet orbital period (P\_P) via the 3$^{th}$ Kepler law (a corollary of Newton's law of universal gravitation)
\item\textbf{Mass and Radius}: the (estimated) planet mass (P\_M) and its radius (P\_R) were considered (these are also useful to distinguish the super-earth planets) 
\item\textbf{Temperature}: the planet equilibrium temperature (P\_T\_E) defined according to the expression $T_{eq}=T_{star}\sqrt{R/2a}\left(1-A\right)^{0.25}$) where R is the star radius (S\_R), a the planet mean distance (P\_D) and A the albedo here considered as 0.3 was considered as well the planet mean stellar flux $P_F$ 
\item\textbf{Habitability}: The planet habitability was classified with a boolean variable using the values reported in the dataset \cite{planet_dataset}

@@

@@

>>>>

\documentclass[compress]{beamer}
\usepackage{hyperref}
\hypersetup{
    pdftitle = {A bird’s-eye view on the habitability of exoplanets via statistical learning techniques}, % <-- Missing in PDF
    pdfauthor = {Marzio De Corato},
    pdfsubject = {Marzio De Corato project the exam: Machine learning, statistical
learning, deep learning and artificial intelligence - Unsupervised Learning}
}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\author{Marzio De Corato}
\usepackage{varwidth}
\usepackage{beamerthemesplit}

@@

\usebackgroundtemplate{\includegraphics[width=\paperwidth]{Pic/Intro.jpg}}
\title[Exoplanets survey via SL techniques]{A bird’s-eye view on the habitability of exoplanets via statistical learning techniques}

\author{}
\date{}

@@

\setbeamertemplate{enumerate items}[default]
\usecolortheme{beaver}

@@

\frame{\vspace{+4.5cm}\titlepage}

\usebackgroundtemplate{ }

@@

\begin{frame}
\frametitle{Theoretical background - Planet features }
\begin{itemize}
\item\textbf{Distance}: in this work the mean planet distance from the host star (P\_D), the periastron (P\_PN) and the apastron (P\_A) as well the termal effective distance (P\_D\_E) from the host star  were considered. These quantities constrain the planet orbital period (P\_P) via the 3$^{th}$ Kepler law (a corollary of Newton's law of universal gravitation)
\item\textbf{Mass and Radius}: the (estimated) planet mass (P\_M) and its radius (P\_R) were considered (these are also useful to distinguish the super-earth planets) 
\item\textbf{Temperature}: the planet mean stellar flux $P_F$  and the planet equilibrium temperature (P\_T\_E) were considered. The latter is defined as $T_{eq}=T_{star}\sqrt{R/2a}\left(1-A\right)^{0.25}$) where R is the star radius (S\_R), a the planet mean distance (P\_D), A the albedo (here it was set equal to 0.3)
\item\textbf{Habitability}: The planet habitability was classified with a boolean variable using the values reported in the dataset \cite{planet_dataset}

@@

\section{Results}
\subsection{Results - Decision Tree}
\begin{frame}
\frametitle{Results - Decision Tree}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.8\linewidth]{Pic/DecisionTree.pdf}
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\linewidth]{Pic/Var_imp_dec_tree.pdf}
\includegraphics[width=0.5\linewidth]{Pic/Decision_Tree_FINAL_ROC.pdf}\\
$\phi_{DT}=0.790$
\end{center}
\end{columns}
\end{frame}

\subsection{Results - Random Forest}
\begin{frame}
\frametitle{Results - Random Forest}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=\linewidth]{Pic/Proximity_Random_forest.pdf}
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.6\linewidth]{Pic/Random_forest_importance.pdf}
\includegraphics[width=0.6\linewidth]{Pic/Random_Forest_FINAL_ROC.pdf}\\
$\phi_{RF}=0.835$
\end{center}
\end{columns}
\end{frame}

\subsection{Results - SVC}
\begin{frame}
\frametitle{Results - Support Vector Classifier}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=\linewidth]{Pic/PCA+SVM_PLOT.pdf}
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\linewidth]{Pic/SVM_FINAL_ROC.pdf}
\includegraphics[width=0.5\linewidth]{Pic/SVM_PCA_ROC.pdf}\\
$\phi_{SVC}=0.527$\quad$\phi_{PCA+SVC}=0.667$
\end{center}
\end{columns}
\end{frame}

\subsection{Results - Logistic Classification}
\begin{frame}
\frametitle{Results - Logistic Classification}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=\linewidth]{Pic/LOG_FULL.png}
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.85\linewidth]{Pic/LOG_FULL_ROC.pdf}\\
\end{center}
\end{columns}
\begin{center}
$\phi_{log}=0.566$
\end{center}
\end{frame}

\subsection{Results -  LDA and QDA}
\begin{frame}
\frametitle{Results - Linear and quadratic discriminant analysis}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=\linewidth]{Pic/QDA_project_FULL.pdf}
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\linewidth]{Pic/LDA_FULL_ROC.pdf}
\includegraphics[width=0.5\linewidth]{Pic/QDA_FULL_ROC.pdf}\\
$\phi_{LDA}=0.526$\quad$\phi_{QDA}=0.757$
\end{center}
\end{columns}
\end{frame}

\subsection{Discussion}
\begin{frame}
\frametitle{Discussion}
\begin{table}[]
\begin{center}
\begin{tabular}{c||c|c|c}
{\color[HTML]{CB0000} \textbf{Algortim}} & {\color[HTML]{CB0000} \textbf{Accuracy}} & {\color[HTML]{CB0000} \textbf{NIR}} & {\color[HTML]{CB0000} \textbf{$\phi$}} \\ \hline\hline
Decision tree                            & 0.952                                    & 0.867                               & 0.790                                  \\ \hline
Random forest                            & 0.964                                    & 0.869                               & 0.835                                  \\ \hline
SVM                                      & 0.907                                    & 0.869                               & 0.527                                  \\ \hline
PCA+SVM                                  & 0.932                                    & 0.869                               & 0.667                                  \\ \hline
Logistic                                 & 0.914                                    & 0.869                               & 0.566                                  \\ \hline
LDA                                      & 0.902                                    & 0.898                               & 0.526                                  \\ \hline
QDA                                      & 0.941                                    & 0.869                               & 0.757

\end{tabular}
\end{center}
\end{table}
\begin{itemize}
\item The random forest algorithm provides the best forecast, followed by the decision tree and the the QDA. 
\item This behaviour can be explained by considering that in the features space,as showed in the previous plots, the two sets largely overlaps
\item A similar performance ranking was found by Sasha et al. (2018) \cite{saha2018machine}

\begin{frame}
\begin{center}
\includegraphics[width=0.5\linewidth]{Pic/Exoplanet_picture.jpg}
\end{center}
\begin{center}
The 2M1207b exoplanet (red spot) and its host star 2M120 (a brown dwarf) as observed, after three near-infrared exposures (H, K and L wavebands), by the VLT Yepun telescope at the ESO Paranal Observatory. Image taken from \cite{exoplanet_pic}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\textit{Per aspera ad astra}
\end{center}
\end{frame}

@@

\section{Supporting slides}
\begin{frame}
\begin{center}
SUPPORTING SLIDES
\end{center}
\end{frame}

\begin{frame}
\frametitle{Proximity plot - Explanation}
\begin{center}
\includegraphics[width=0.8\linewidth]{Pic/Prox_explanation.jpg}\\
Image taken from \cite{blanchet2020constructing}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\frametitle{Proximity plot - Explanation}
\includegraphics[width=0.8\linewidth]{Pic/Prox_explanation_3.jpg}\\
Image taken from \cite{blanchet2020constructing}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\frametitle{Proximity plot - Explanation}
\includegraphics[width=0.8\linewidth]{Pic/Prox_explanation_2.jpg}\\
Image taken from \cite{blanchet2020constructing}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\frametitle{$\phi$ coefficent}
\begin{table}[]
\begin{tabular}{c|c|c}
\multicolumn{1}{l|}{\textbf{}} & \textbf{Actual - N} & \textbf{Actual - P} \\ \hline
\textbf{Predicted - N}         & \#TP                & \#FN                \\ \hline
\textbf{Predicted - P}         & \#FP                & \#TP               
\end{tabular}
\end{table}
\begin{equation*}
\phi=\dfrac{TP\times TN - FP\times FN}{\sqrt{\left(TP+FP\right)\left(TP+FN\right)\left(TN+FP\right)\left(TN+FN\right)}}
\end{equation*}
\end{center}
\end{frame}
